---
title: "Initial Korean Covid 19"
author: "Lizeth Marquez"
date: "30 May 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```



#Introduction



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these  
state of a patient must be died or alive - we adjust that as well       
  
# Your Question - Is dying independent of sex?

$H_0$:
There is no difference in death rates depending upon on sex.


$H_A$:
There is a difference in death rates depending upon on sex.

# Methods

The result is a categorical variable, sex is categorical.   
The result depends on the sex, we use the helper file categorical~categorical.


## Graphical results - use ggplot commands to do this  
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The bar chart shows that there is a higher percentage of females than males alive.   
It also reflects that there is a higher percentage of males than women dead.


  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)
```
This numerical result shows that out of 100 people, 64 males and 36 females died.   
It resulted in 3% of males dead, and 3% of women.  

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The p-value for the chi-squared test is .001, so we must reject the null hypothesis.   
Based on the inferential results, the sex is not related to covid deaths. 

The Fisher's exact test shows that ratio of men are 2.3 times more likely to die than women in Korea due to Covid 19.


## Conclusions

The dying is not dependent on sex origin.   
We reject the null $H_0$: because the p value is lower than .05.   
The data from the inferential results determines that death and gender are not independently correlated. 
